---
title: "DenVar"
#author: Souvik Seal
output: github_document
bibliography: [packages.bib, ref.bib]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#",
  out.width = "100%",
  messages = FALSE,
  warnings = FALSE
)
```

This is a R package implementing the proposed method from the paper "DenVar: Density-based Variation analysis of multiplex imaging data".

## Loading required packages

First, we load our package, DenVar and a few other required packages. One can install the developmental version of DenVar by running the command: **devtools::install_github('sealx017/DenVar')**. 


```{r loading packages, message=FALSE}
#devtools::install_github('sealx017/DenVar')
require(DenVar)
require(LaplacesDemon)
require(pheatmap)
require(ggplot2)
require(survival)
require(survminer)
require(coxme)

```

## Loading the dataset 

Next, we import the example datasets named, "PD1_TNBC_MIBI.csv" and "Survival_TNBC_MIBI.csv" which are respectively the PD-1 marker expression data and the survival data of 33 patients from TMBC MIBI data. Both the datasets have a column named "SampleID" denoting individual patient IDs.

```{r loading the marker expression and the clinical data}

PD1_data = read.csv("Data/PD1_TNBC_MIBI.csv")[,-1]
clinical_data = read.csv("Data/Survival_TNBC_MIBI.csv")[,-1]

```


## Computing the KDEs and the JSD matrix

We compute the KDE of PD-1 expression in all the patients and store it in an array. We keep number of grid points (ngrids) in the KDE at 1024. Using the array of KDEs, we compute pairwise Jensen-Shannon Distance (JSD) between the patient densities. We create the JSD matrix having distances between all possible patient pairs.

```{r Computing the KDEs and the JSD matrix}

KDE_of_all = Array_KDE(Data = PD1_data, ngrids = 1024)
JSD_matrix_between_all = JSD_matrix(Array_dens = KDE_of_all, Data = PD1_data)

```
## Heatmap of the JSD matrix

Next, we visualize the computed JSD matrix as a heatmap. 

```{r heatmap of JSD matrix}

print(pheatmap(JSD_matrix_between_all, cluster_rows = TRUE,
cluster_cols = TRUE,show_colnames = T, show_rownames = T,treeheight_row=0,treeheight_col = 0))

```

## Hierarchical clustering based on the JSD matrix

We use the JSD matrix for hierarchical clusetring of the patients into two meanigful clusters.

```{r hclust on the JSD matrix}

hier_clustering = hclust(as.dist(JSD_matrix_between_all), method="ward.D")
hier_groups = as.data.frame(cutree(hier_clustering, k=2)); colnames(hier_groups) = "Cluster"

```

## CoxPH model and KM plots

Using the vector of cluster-labels, we fit a CoxPH model and plot the Kaplan-Meyer (KM) curves stratified by the clusters. We display the estimated Hazard Ratio (HR) and the corresponding p-value as well.

```{r CoxPh model with the clusters}

surv = cbind(clinical_data, hier_groups)
coxplot = coxPH_plot(surv)
#print(coxplot$plot)

```

## CoxPH model with random effects (Coxme)

We use the computed JSD matrix directly in a CoxPH model with random effect (also known as Coxme). We show the estimated variance component and the corresponding LRT for testing its significance.


```{r Coxme model with the clusters}

res.coxme = coxme_model(surv, JSD_matrix_between_all)
print(res.coxme)

```
```{r, include=FALSE}
knitr::write_bib(file = 'packages.bib')
```
